Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean in the United States

### 2017-08-15 doc/manuscript/manuscript.Rmd
Add McCoy as Author; update DNA isolation protocol

---
title: Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean in the United States
preprint: false
author:
  - name: Zhian N. Kamvar
    email: zkamvar@unl.edu
    affiliation: 1
  - name: Bimal Sajeewa Amadarasa
    email: bamadarasa2@unl.edu
    affiliation: 1
  - name: Rachana Jhala
    email: rjhala2@unl.edu
    affiliation: 1
  - name: James Steadman
    email: jsteadman1@unl.edu
    affiliation: 1
  - name: Sydney E. Everhart
    email: everhart@unl.edu
    affiliation: 1
    corresponding: true
affiliation:
  - code: 1
    address: Department of Plant Pathology, University of Nebraska-Lincoln, Lincoln 68583
abstract: >
  `r paste(readLines("abstract.md"), collapse = "\n  ")`
header-includes: >
  \usepackage{indentfirst}
bibliography: ssc_bibliography.bib
output: 
  rticles::peerj_article:
    citation_package: natbib
    base_format: bookdown::pdf_document2 # for using \@ref()
---

@@

Each isolate was regenerated by activating one or several sclerotia and transferring the leading edge of the mycelial mat according to the method described previously [@otto-hanson2011variation].
In combination with the 156 isolates described previously, we collected 210 isolates for a total of 366 isolates [@otto-hanson2011variation].

@@

@@

> SEE: More details on how tissue was disrupted may be needed here.

@@

Of interest in this study was testing the association between multilocus genotypes (MLGs) and mycelial compatibility groups (MCGs), as the latter have historically been used as a proxy for clonal lineages [@leslie1993fungal; @otto-hanson2011variation; @kohn1990mycelial; @schafer2006optimized; @kohn1995clonal; @carbone1999patterns]. 
Our results do not support this assumption.
It can be seen in Fig. \@ref(biggraph)A that the most abundant MCG contains several MLGs, but the diversity of those MLGs are low as indicated by the evenness (transparency), which indicates that there is one dominant MLG (MLG 25).
What is not shown in Fig. \@ref(biggraph)A is the MLGs that are shared between MCGs. 
This is illustrated in both table \@ref(tab:mlg-table) and Fig. \@ref(biggraph)B,C. 
It could be argued, however that MLG 25, with its high value of $P_{sex}$ represents different true MLGs across the five MCGs it occupies, but this doesn't account for the overall structure of Fig. \@ref(fullgraph) where, for example, MLG 75 ($P_{sex}$ = 1.81e^-4^) is compatible with 58 other genotypes through three MCG when the population structure of *S. sclerotiorum* is known to be clonal.

@@

The results show several genotypes shared across populations, except for Mexico and two states that had fewer than three samples each (Fig. \@ref(community-graph)).
Our network-based approach by treating the genotypes as edges and weighting each edge with the inverse of $P_{sex}$ treated the edges as springs connecting the populations with the strength proportional to the probability of obtaining the same genotype as a clone. 
This allowed us to use a graph walking algorithm to see how close the populations were simply based off of the proportion of clones they shared.
The most abundant genotype was shared across four populations, but its high value of $P_{sex}$ meant that it did not contribute significantly to the overall structure. 
The graph walking algorithm was able to divide the network into three groups, but had a modularity of 0.17, which indicates that the groups are weakly differentiated.
This indicates that multilocus genotypes are widespread across the United States, giving more evidence for a clonal population structure.

>>>>

---
title: Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean in the United States
preprint: false
author:
  - name: Zhian N. Kamvar
    email: zkamvar@unl.edu
    affiliation: 1
  - name: Bimal Sajeewa Amadarasa
    email: bamadarasa2@unl.edu
    affiliation: 1
  - name: Rachana Jhala
    email: rjhala2@unl.edu
    affiliation: 1
  - name: Serena McCoy
    email: ¯\_(ツ)_/¯
    affiliation: 1
  - name: James Steadman
    email: jsteadman1@unl.edu
    affiliation: 1
  - name: Sydney E. Everhart
    email: everhart@unl.edu
    affiliation: 1
    corresponding: true
affiliation:
  - code: 1
    address: Department of Plant Pathology, University of Nebraska-Lincoln, Lincoln 68583
abstract: >
  `r paste(readLines("abstract.md"), collapse = "\n  ")`
header-includes: >
  \usepackage{indentfirst}
bibliography: ssc_bibliography.bib
output: 
  rticles::peerj_article:
    citation_package: natbib
    base_format: bookdown::pdf_document2 # for using \@ref()
---

@@

After collection, the isolates were stored in petri plates lined with filter paper, then stored at $20^{\circ}$F or –$4^{\circ}$C. 
Sclerotia were triple rinsed with (i) 50% Clorox bleach/50% double-distilled (dd)H2O solution for 3 min, (ii) ddH2O rinse for 3 min, and (iii) final ddH2O rinse for 3 min, then plated on water agar plates (16g of Bacto agar per liter of ddH2O), with four to five sclerotia of each isolate separated on each plate and stored on the bench top at room temperature for 5 to 6 days An 8-mm plug from a 5- or 6-day-old culture was transferred from the advancing margin of the mycelia onto a plate of Difco potato dextrose agar (PDA at 39 g/liter of ddH2O) [@otto-hanson2011variation].
In combination with the 156 isolates described previously, we collected 210 isolates for a total of 366 isolates [@otto-hanson2011variation].

@@

 <!-- Preceeding the steps in that, we grew the isolate(s) on PDA, and transferred plugs to Potato Dextrose Broth. We grew the culture in broth until there was significant mycelial growth, but before the mycelial mat became 'solid' (about 4-5 days).  We collected the mycelia using a buchner funnel, removed the agar plugs, lyophilized the mycelia, then pulverized it by hand in whirl-pak bags.  The lyophilized mycelia was then used in the procedure as printed. Lyophilzed mycelia was stored in microcentrifuge tubes in the -20 freezer until needed for extractions.  -->

Prior to DNA extraction, isolates were grown on PDA and plugs were subsequently transferred to Potato Dextrose Broth (PDB) where they were grown until there was significant mycelial growth, but before the mycelial mat became solidified (4--5 days).
Mycelial was collected in a buchner funnel, agar plugs removed, lyophilizedm and pulverized manually in whirl-pak bags. 
Lyophilized mycelia was then stored in microcentrifuge tubes at -$20^{\circ}$C until needed for DNA extraction.

@@

<!-- Fungal DNA mini-prep Updated from Jeff Rollins’ protocol           Rev. Nov 2012

Sclerotinia spp. Phenol/Chloroform extraction-

1) Spin 2.0 ml Phase-lock gel microcentrifuge tubes @14,000 rpm for 30 sec.- 2 for each sample to be extracted

2) Place 600 ul LETS buffer in 2.0 ml microcentrifuge (round-bottomed)  tube

3) Add ~25 mg of lyophilized mycelia; mix by inverting (shaking, flicking) tube.

4) Add an equal volume (600ul) of buffer-saturated Phenol.  Invert to mix.

5) Add 300ul chloroform to tubes- invert to mix.

6) Pour suspension into Phase-lock  gel tubes.

7) Spin Phase-lock tubes @12,000 RPM for 8-10 min. to separate phases.

8) Add another volume (300ul per tube) of chloroform, invert gently to mix.  
    Spin again @ 12,000 RPM for 8 min.

9) Pour off top (aqueous) layer into micro-centrifuge tube.

10) Add an equal volume 2-propanol to volume in each tube.  Mix by gently inverting. Nucleic acids will be visible as transluscent strands in tubes.  
(Optional- Place in freezer (-20OC) overnight.)

**Turn on heat block to ‘low’ setting**

11) Spin tubes @ 12,000 RPM for 14 min.  White or loose semi-translucent pellets should be visible in tube.

12) Rinse pellet with ice-cold 70% EtOH, pouring off alcohol carefully- watch out for ‘loose’ pellets (re-spin if necessary).  Invert on paper towel for ~ 20 min to dry.

13) Resuspend pellet in 450 ul TE, then add 10 ul RNAase to each tube- invert or flick to dissolve pellet.  Incubate tubes on heat block at 37OC for 30 min.

14) Add 450 ul Phenol to each tube.  Invert tubes to mix.

16) Add 450 ul chloroform to each tube, invert to mix.

15) Transfer tube contents to Phase-lock  tubes, then spin @ 14,000 RPM for 10 min. to separate phases

16) Add 300 ul chloroform to each tube, invert to mix.

17) Spin @ 14,000 RPM for 8-10 min. to again separate phases.

18) Using wide-bore pipette tip, transfer aqueous phase to 1.5 micro-centrifuge tube- recover up to 450ul (not more) to keep total volume within capacity of micro-centrifuge tube.

19) Add 0.1 volume (45ul) 3M NaOAc to each tube.  Add two volumes ice-cold absolute EtOH (fill tube).  Incubate on ice for 30 min.- translucent material should be visible in tubes.
(Optional- Place in freezer (-20OC) overnight.)

20) Centrifuge at 14,000 RPM for 10 min.

21) Pour off EtOH, being careful not to dislodge pellet.  Rinse in cold 70% EtOH, then dry pellet on paper towels for at least 15 min.  Pellets may appear as white “pellets” in bottom of tube, or may appear as translucent smears on side of tube.

22) Resuspend pellets in 200ul TE.  (Place in 4OC refrig for ~24 hrs to allow pellet to rehydrate). -->

@@

Of interest in this study was testing the association between multilocus genotypes (MLGs) and mycelial compatibility groups (MCGs), as the latter have historically been used as a proxy for clonal lineages [@leslie1993fungal; @otto-hanson2011variation; @kohn1990mycelial; @schafer2006optimized; @kohn1995clonal; @carbone1999patterns]. 
Our results do not support this assumption.
It can be seen in Fig. \@ref(biggraph)A that the most abundant MCG contains several MLGs, but the diversity of those MLGs are low as indicated by the evenness (transparency), which indicates that there is one dominant MLG ('MLG 25').
What is not shown in Fig. \@ref(biggraph)A is the MLGs that are shared between MCGs. 
This is illustrated in both table \@ref(tab:mlg-table) and Fig. \@ref(biggraph)B,C. 
It could be argued, however that 'MLG 25', with its high value of $P_{sex}$ represents different true MLGs across the five MCGs it occupies, but this doesn't account for the overall structure of Fig. \@ref(fullgraph) where, for example, 'MLG 75' ($P_{sex}$ = 1.81e^-4^) is compatible with 58 other genotypes through three MCG when the population structure of *S. sclerotiorum* is known to be clonal.

@@

The results show several genotypes shared across populations, except for Mexico and two states that had fewer than three samples each (Fig. \@ref(community-graph)).
Our network-based approach by treating the genotypes as edges and weighting each edge with the inverse of $P_{sex}$ treated the edges as springs connecting the populations with the strength proportional to the probability of obtaining the same genotype as a clone. 
This allowed us to use a graph walking algorithm to see how close the populations were simply based off of the proportion of clones they shared.
The most abundant genotype was shared across four populations, but its high value of $P_{sex}$ meant that it did not contribute significantly to the overall structure. 
The graph walking algorithm was able to divide the network into three groups, but had a modularity of 0.17, which indicates that the groups are only weakly differentiated.
This indicates that multilocus genotypes are widespread across the United States, giving more evidence for a clonal population structure.
